A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][Tf_2N]}

### 2019-02-12 ionic_liquids.tex
Submitted version
revision I

\usepackage[colorlinks,citecolor=blue]{hyperref}
%\usepackage{lineno}
%\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{natbib}
\usepackage[version=4]{mhchem}
\usepackage{adjustbox}   
\usepackage[position=top,captionskip=0pt]{subfig}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\usepackage{xcolor}
\usepackage{pdfcomment}

@@

\title{A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}

@@

\begin{abstract}
In this paper, we calculate solvation free energies of several compounds in ionic liquids.
These free energies are used to compute properties such as Henry's law constants and activity coefficients, generally required in the design of environmentally sustainable processes.
It is known, for instance, that carbon dioxide from combustion is one of the main sources of anthropogenic greenhouse gases.
Recently, the propensity of the ionic liquid \ce{[emim][B(CN)_4]} for the physisorption of \ce{CO_2} has been reported, which makes it a potential solvent for carbon capture.
In the present work, molecular dynamics simulations of the solvation of \ce{CO_2} in ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} are carried out both at infinite dilution and at high concentrations.
A systematic study is performed by comparing several force fields and assessing the efficacy of simplification measures such as rigid-body dynamics and pairwise electrostatics.
Our results confirm recent experimental observations that, for a given volume of solvent, lower pressure is required to absorb a certain amount of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in other ionic liquids.
\end{abstract}

@@

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures.
The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA).
Despite its high absorption capability, the use of MEA has a serious drawback, which is the high cost of solvent recovery \cite{Merkel_2010}.
Other disadvantages include corrosion and degradation by oxidation, as well as solvent loss by evaporation.
The latter also contributes to the increase of environmental pollution and generates extra costs for replacing the solvent.

@@

Recently, Carvalho \textit{et al}. \cite{Carvalho_2016} called into question the potentiality of ILs for \ce{CO_2} capture.
They warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposedly high absorption capability of certain ILs.
Although most thermodynamic models are formulated in terms of molar fractions, a key factor that dictates the efficiency of absorption process  is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent.
By expressing the solubility values in terms of molality and performing an exhaustive comparison among various ILs, Carvalho \textit{et al}. \cite{Carvalho_2016} show that most systems which promote physical absorption of \ce{CO_2} exhibit solubilities which are similar, or even lower, than that observed for heavy alkanes such as eicosane.
This is true for the popular ILs \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}.
The only exception they found was the family of ILs containing the anion \ce{[B(CN)_4]^-} (tetracyanoborate).

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of \ce{[B(CN)_4]^-}.
In this context, molecular simulation provides us with a powerful tool for studying solvation processes.
To the best of our knowledge, only one study in the literature reports computation of solvation free energy for \ce{CO_2} in \ce{[emim][B(CN)_4]} by molecular simulation \cite{Liu_2014_1}.
Nonetheless, the computed value is overestimated, as observed by deriving a Henry's law constant and comparing it to available experimental data.
Such scarcity motivated us to carry out the present investigation, in which we have evaluated different force fields for \ce{[emim][B(CN)_4]} proposed in the literature.
As we show here, it was possible to find a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2} in this IL.
For comparison, we analyzed the solvation of \ce{CO_2} not only in \ce{[emim][B(CN)_4]}, but also in \ce{[emim][NTf_2]}, both at infinite dilution and at high concentrations.
The simulations confirmed what is observed in experiments.
For \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}, considering a given mass of solvent.
In addition, we also calculated infinite-dilution activity coefficients of other solutes in \ce{[emim][B(CN)_4]} and compared the results with experimental values recently reported in literature.
The solutes we considered were hexane, benzene, cyclohexane, ethanol, and water.

@@

In what follows, the subscripts ``s'' and ``IL'' refer to the solute and the solvent (ionic liquid), respectively.
For a liquid phase containing $N_\text{IL}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_\text{sim} \equiv R T \ln \frac{\Xi (N_s,N_\text{IL},P,T,\lambda = 1)}{\Xi (N_s,N_\text{IL},P,T,\lambda = 0)}, 
\end{equation}
where $\Xi$ denotes the isobaric-isothermal partition function and $R$ is the gas constant.
Shirts \textit{et al}. \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_\text{sim}$ to what is experimentally measured as the solvation free energy, namely
\begin{equation}
\label{eq:solvation free energy definition}
\Delta G_{\text{solv}} \equiv R T \ln \left( \frac{P_s}{c_s R T} \right),
\end{equation}
where $P_s$ is the partial pressure of the solute in the gas phase and
$c_s$ is the molar concentration of the solute in the liquid phase.
Note that these quantities correspond to a phase equilibrium condition.
For obtaining the mentioned relation, the authors derived an expression for the chemical potential of the solute in the liquid phase ($\mu^l_s$) as a function of $\Delta G_\text{sim}$, which can be written as
\begin{equation}
\label{eq:solute chemical potential}
\mu^l_s = \mu^{ig}_s + \Delta G_\text{sim} - RT \ln \left( \frac{V^{\ast}}{V_1}\right) - RT \ln \left( \frac{P_s}{c_s R T} \right),
\end{equation}
where $\mu^{ig}_s$ is the chemical potential of the solute in an ideal gas at temperature $T$ and solute partial pressure $P_s$, while $V^{\ast}$ and $V_1$ are the mean volumes of liquid solutions specified by ($N_s-1$,$N_\text{IL}$,P,T) and ($N_s$,$N_\text{IL}$,P,T), respectively.
Note that $c_s = \frac{N_s}{N_A V_1}$, where $N_A$ is the Avogadro number.

@@

If both $P_s$ and $c_s$ are small, we can resort to Eqs.~\eqref{eq:solvation free energy definition} and \eqref{eq:free_solv} to predict Henry's law constant, which is defined as $\text{K}_s =\frac{P_s}{x_s}$ \cite{Prausnitz}.
%\begin{equation}
%\text{K}_s =\frac{P_s}{x_s}.
%\end{equation}
%
Therefore, by simply expressing Eq.~\eqref{eq:solvation free energy definition} in terms of the molar fraction of the solute in the liquid phase ($x_s$), we can conclude that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_\text{sim} R \, T \exp\left(\frac{\Delta G_\text{sim}}{R T}\right),
\end{equation}
where $\tilde{\rho}_\text{sim}$ is the simulated molar density of the solution. Assuming that this solution is very dilute, one can use the molar density of pure solvent ($\tilde{\rho}_\text{IL}$) instead.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} in the liquid phase are obtained at pressure values for which, in principle, the ideal gas assumption might no longer be valid.
In this case, we can compute the chemical potential of the solute in the real gas phase as
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + R T \ln \left(\frac{f^{g}_s}{P_s}\right),
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase.
Then, by considering a system at phase equilibrium ($\mu^{l}_s$ = $\mu^{g}_s$), it follows that
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = c_s R T \exp \left( \frac{\Delta G_\text{sim}}{R T} \right). 
\end{equation}

Three additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems with which we compare our simulations) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equilibrium pressure}
y_s \exp\left( \frac{B_s P}{R T} \right) P = c_s R T \exp \left( \frac{\Delta G_\text{sim}}{R T} \right),
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

Our third assumption, supported by preliminary simulations, is that $\Delta G_\text{sim}$ does not vary appreciably with pressure, at least over the range of experimental data considered here.
In this way, no iterative simulations are required to compute the equilibrium pressure when solving Eq.~\eqref{eq:equilibrium pressure} for $P$.

@@

As will be shown below, the prediction of activity coefficients requires additional simulations of the solute as a pure liquid phase, whose properties we denote here by a superscript ``$0$''.
We start by using a standard thermodynamic relation \cite{Tester} to write the solute activity coefficient $\gamma_s$ in terms of chemical potentials as
\begin{equation}
\ln \gamma_s = \frac{\mu_s^l - \mu_s^{l,0}}{R T} - \ln x_s,
\end{equation}

@@

In the special case of a solute at infinite dilution in an ionic liquid solvent, its activity coefficient $\gamma^\infty_s$ can be computed by replacing $\tilde{\rho}$ in the equation above by the molar density of the pure solvent, $\tilde{\rho}_\textit{IL}$.

@@

Finally, we also consider the parameterization which Weber and Kirchner \cite{Weber_2016} carried out for \ce{[emim][B(CN)_4]} based on the force field developed by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
As the potential parameters for the anion were specified neither in Ref.~\citenum{Weber_2016} nor in Ref.~\citenum{Canongia_Lopes_2006}, we obtained them by direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure with Hartree-Fock calculations at the \mbox{6-31++G\textsuperscript{**}} level of theory, resulting in net charges of $\pm 0.8 ~\mathrm{e}$.

In the case of ionic liquid \ce{[emim][NTf_2]}, we use the model of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007}, which was also derived by refining parameters of the force field by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
The authors characterized this refinement as unusual, given that they modified the LJ parameters while keeping the partial charges unchanged.
Kerl \textit{et al}. \cite{Kerl__2017} recently employed this model to accurately predict Henry constants of \ce{CO_2} in \ce{[emim][NTf_2]} at several temperatures.

@@

With the aim of setting up a computationally efficient protocol for the IL simulations, in this work we assess the possibilities of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing electrostatic interactions.
In this method, the electrostatic potential between two atoms with charges $q_i$ and $q_j$ interacting at a distance $r$ is given by
\begin{equation}
V_\text{DSF}(r) = V(r) - V(r_c) - V^\prime(r_c)(r-r_c),
\end{equation}
where $V(r) = q_i q_j {\text{erfc}(\alpha r)}/{r}$ is a damped Coulomb potential.
This potential is cut off at $r = r_c$, which makes it short-ranged and thus a cheaper alternative to Ewald-type schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}.
McCann \textit{et al}. \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force field.
They concluded that using $\alpha = 0.2~\text{\AA}^{-1}$ and $r_c = 15~\text{\AA}$ provides the best energetic agreement with the Ewald method.
In this work, detailed comparisons are made in terms of the simulated densities and radial distribution functions.
In Sec.~\ref{sec:results}, we show that by using the same $\alpha$ value but a considerably smaller cutoff distance ($r_c = 12 \text{\AA}$) one can obtain suitable results.
These values were employed here for the ILs as well as for solutes containing partially charged atoms.
The only exception was hexane, for which the use of undamped Coulombic interactions (i.e. $\alpha = 0$) resulted in improved chemical potential predictions.

The NPT algorithm of Kamberaj \textit{et al}. \cite{Kamberaj_2005} was used to integrate the equations of motion for rigid bodies.
For flexible molecules, we applied the integration scheme of Shinoda \textit{et al}. \cite{Shinoda2004}, with the lengths of all C-H bonds constrained by the SHAKE method \cite{Ryckaert1977}.
For both integrators, we used a small time-step size ($\Delta t$) of 1 fs in order to avoid discretization-induced artifacts \cite{Eastwood_2010, Silveira_2017}.

@@

In this work, we compute $\Delta G_\text{Coul}$ by carrying out an independent simulation of 10 ns at each state of $\lambda_{\text{Coul}}$.
In the end, we apply the Multistate Bennett Acceptance Ratio (MBAR) method \cite{Shirts_2008} by using the \textit{Alchemical Analysis} package \cite{Klimovich_2015}.

In the case of $\Delta G_\text{vdW}$, the Expanded Ensemble method \cite{Lyubartsev_1992} is applied for sampling the whole $\lambda_\text{vdW}$ set in a single simulation.
An expanded ensemble is constructed as a weighted sum of subensembles, each one assuming a different value of the parameter selected for exploring the free energy landscape, which in our case is $\lambda_\text{vdW}$.
In the method adopted here \cite{Chodera_2011_2}, simulating an expanded ensemble comprises doing molecular dynamics (MD) at fixed N, P, T, and $\lambda_\text{vdW}$ amid Monte Carlo (MC) transitions in the value of $\lambda_\text{vdW}$.
The number of MD steps between transitions is fixed and, at each transition, one of the $\lambda_\text{vdW}$ states is picked at random according to a corresponding set of probabilities \cite{Chodera_2011_2}.
In addition, a set of biasing weights is included in the MC acceptance criterion, thus making it possible to enhance the sampling procedure by iteratively tuning these weights until all $\lambda_\text{vdW}$ states are visited with equal frequency \cite{Lyubartsev_1992}.
Given that sample decorrelation is improved by having short MD runs between MC transitions \cite{Chodera_2011_2}, we used only 5 steps in each run.
In the iteration process for tuning the biasing weights, the simulation times for the preliminary and the final expanded ensemble rounds are 10 and 40 ns, respectively.
In this way, 3 iterations were sufficient in all cases for attaining, in the final round, an acceptable uncertainty in $\Delta G_\text{sim}$ estimated via MBAR \cite{Shirts_2008, Klimovich_2015}.

We first carried out MD simulations of isolated ion pairs in order to obtain minimum-energy conformations.
For this, we used a Nos\'{e}-Hoover chain thermostat \cite{Martyna1992} with set-point temperature gradually changing down to nearly $0~\text{K}$.
The resulting coordinates were then used with the PACKMOL algorithm \cite{Mart_nez_2009} via the Playmol package \cite{playmol} to obtain initial liquid-phase configurations.
The expanded ensemble procedure, the softcore potential, and the DSF method  with linear coupling were implemented by customizing the LAMMPS package \cite{Plimpton1995}, which we used for carrying out all the simulations in this work.

@@

\subfloat[Liu~\textit{et al}.~\cite{Liu_2014}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Liu}}%
\subfloat[Weber and Kirchner~\cite{Weber_2016}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Weber}}%
\caption{Combined distribution functions of the dihedral angles corresponding to atoms C1-N2-C5-C6 and C2-N2-C5-C6 in the cation \ce{[emim]^+} (see Fig.~\ref{fig:atoms_id}), as sampled in NPT-MD simulations of 250 ion pairs using each model for \ce{[emim][B(CN)_4]}.}
\label{fig:dihedrals_emim}
\end{figure*}

>>>>

\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{natbib}
\usepackage[version=4]{mhchem}
\usepackage{adjustbox}   
\usepackage[position=top,captionskip=0pt]{subfig}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\usepackage{xcolor}

\usepackage{pdfcomment}

@@

\title{A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][Tf_2N]}}

@@

\begin{abstract}
In this paper, we calculate solvation free energies of several compounds in ionic liquids.
These free energies are used to compute properties such as Henry's law constants and activity coefficients, generally required in the design of environmentally sustainable processes.
It is known, for instance, that carbon dioxide from combustion is one of the main sources of anthropogenic greenhouse gases.
Recently, the propensity of the ionic liquid 1-ethyl-3-methylimidazolium tetracyanoborate (\ce{[emim][B(CN)_4]}) for the physisorption of \ce{CO_2} has been reported, which makes it a potential solvent for carbon capture.
In the present work, molecular dynamics simulations of the solvation of \ce{CO_2} in ionic liquids \ce{[emim][B(CN)_4]} and 1-ethyl-3-methylimidazolium bis (trifluoromethylsulfonyl) imide (\ce{[emim][Tf_2N]}) are carried out both at infinite dilution and at high concentrations.
A systematic study is performed by comparing several force fields and assessing the efficacy of simplifications in the simulations by using rigid-body dynamics and pairwise electrostatics.
Our results confirm recent experimental observations that, for a given volume of solvent, lower pressure is required to absorb a certain amount of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in other ionic liquids.
\end{abstract}

@@

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures.
The most commonly used solvents are aqueous solutions of primary and secondary alkanolamines, such as monoethanolamine (MEA).
Despite its high absorption capability, the use of MEA has a serious drawback, which is the high cost of solvent recovery \cite{Merkel_2010}.
Other disadvantages include corrosion and degradation by oxidation, as well as solvent loss by evaporation.
The latter also contributes to the increase of environmental pollution and generates extra costs for replacing the solvent.

@@

Recently, Carvalho \textit{et al}. \cite{Carvalho_2016} called into question the potentiality of ILs for \ce{CO_2} capture.
They warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposedly high absorption capability of certain ILs.
Although most thermodynamic models are formulated in terms of molar fractions, a key factor that dictates the efficiency of absorption process  is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent.
By expressing the solubility values in terms of molality and performing an exhaustive comparison among various ILs, Carvalho \textit{et al}. \cite{Carvalho_2016} show that most systems which promote physical absorption of \ce{CO_2} exhibit solubilities which are similar, or even lower, than that observed for heavy alkanes such as eicosane.
This is true for the popular ILs \ce{[bmim][Tf_2N]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][Tf_2N]} and \ce{[hmim][pFAP]}.
The only exception they found was the family of ILs containing the anion \ce{[B(CN)_4]^-} (tetracyanoborate).

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of \ce{[B(CN)_4]^-}.
In this context, molecular simulation provides us with a powerful tool for studying solvation processes.
To the best of our knowledge, only one study in the literature reports computation of solvation free energy for \ce{CO_2} in \ce{[emim][B(CN)_4]} by molecular simulation \cite{Liu_2014_1}.
Nonetheless, the computed value is overestimated, as observed by deriving a Henry's law constant and comparing it to available experimental data.
Such scarcity motivated us to carry out the present investigation, in which we have evaluated different force fields for \ce{[emim][B(CN)_4]} proposed in the literature.
As we show here, it was possible to find a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2} in this IL.
For comparison, we analyzed the solvation of \ce{CO_2} not only in \ce{[emim][B(CN)_4]}, but also in \ce{[emim][Tf_2N]}, both at infinite dilution and at high concentrations.
The simulations confirmed what is observed in experiments.
For \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}, considering a given mass of solvent.
In addition, we also calculated infinite-dilution activity coefficients of other solutes in \ce{[emim][B(CN)_4]} and compared the results with experimental values recently reported in literature.
The solutes we considered were hexane, benzene, cyclohexane, ethanol, and water.

@@

In what follows, the subscripts ``s'' and ``IL'' refer to the solute and the solvent (ionic liquid), respectively.
For a liquid phase containing $N_\text{IL}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (p) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_\text{sim} \equiv R T \ln \frac{\Xi (N_s,N_\text{IL},p,T,\lambda = 1)}{\Xi (N_s,N_\text{IL},p,T,\lambda = 0)}, 
\end{equation}
where $\Xi$ denotes the isobaric-isothermal partition function and $R$ is the gas constant.
Shirts \textit{et al}. \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_\text{sim}$ to what is experimentally measured as the solvation free energy, namely
\begin{equation}
\label{eq:solvation free energy definition}
\Delta G_{\text{solv}} \equiv R T \ln \left( \frac{p_s}{c_s R T} \right),
\end{equation}
where $p_s$ is the partial pressure of the solute in the gas phase and
$c_s$ is the molar concentration of the solute in the liquid phase.
Note that these quantities correspond to a phase equilibrium condition.
For obtaining the mentioned relation, the authors derived an expression for the chemical potential of the solute in the liquid phase ($\mu^l_s$) as a function of $\Delta G_\text{sim}$, which can be written as
\begin{equation}
\label{eq:solute chemical potential}
\mu^l_s = \mu^{ig}_s + \Delta G_\text{sim} - RT \ln \left( \frac{V^{\ast}}{V_1}\right) - RT \ln \left( \frac{p_s}{c_s R T} \right),
\end{equation}
where $\mu^{ig}_s$ is the chemical potential of the solute in an ideal gas at temperature $T$ and solute partial pressure $p_s$, while $V^{\ast}$ and $V_1$ are the mean volumes of liquid solutions specified by ($N_s-1$,$N_\text{IL}$,p,T) and ($N_s$,$N_\text{IL}$,p,T), respectively.
Note that $c_s = \frac{N_s}{N_A V_1}$, where $N_A$ is the Avogadro number.

@@

If both $p_s$ and $c_s$ are small, we can resort to Eqs.~\eqref{eq:solvation free energy definition} and \eqref{eq:free_solv} to predict Henry's law constant, which is defined as $\text{K}_s =\frac{p_s}{x_s}$ \cite{Prausnitz}.
%\begin{equation}
%\text{K}_s =\frac{p_s}{x_s}.
%\end{equation}
%
Therefore, by simply expressing Eq.~\eqref{eq:solvation free energy definition} in terms of the molar fraction of the solute in the liquid phase ($x_s$), we can conclude that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_\text{sim} R \, T \exp\left(\frac{\Delta G_\text{sim}}{R T}\right),
\end{equation}
where $\tilde{\rho}_\text{sim}$ is the simulated molar density of the solution. Assuming that this solution is very dilute, one can use the molar density of pure solvent ($\tilde{\rho}_\text{IL}$) instead.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} in the liquid phase are obtained at pressure values for which, in principle, the ideal gas assumption might no longer be valid.
In this case, we can compute the chemical potential of the solute in the real gas phase as
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + R T \ln \left(\frac{f^{g}_s}{p_s}\right),
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase.
Then, by considering a system at phase equilibrium ($\mu^{l}_s$ = $\mu^{g}_s$), it follows that
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = c_s R T \exp \left( \frac{\Delta G_\text{sim}}{R T} \right). 
\end{equation}

Two additional assumptions are made here. First, we assume that the gas phase is either pure solute (which is consistent with the ionic-liquid systems investigated) or can be approximated by an ideal solution. Second, we assume that the system pressure is moderate, so that a truncated virial equation is enough to represent the gas phase. In this case, Eq.~\eqref{eq:fgas_d} becomes
\begin{equation}
\label{eq:equilibrium pressure}
y_s \exp\left( \frac{B_s p}{R T} \right) p = c_s(p) R T \exp \left( \frac{\Delta G_\text{sim}}{R T} \right),
\end{equation}
where $y_s$ is the molar fraction of solute in the gas phase ($y_s \rightarrow 1$ for the systems we analyze here) and $B_s$ is the second virial coefficient of the solute.
In the case of \ce{CO_2}, we employ the correlation for $B_s$ developed by Holste \textit{et al}. \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0 = 23.02991~\mathrm{\frac{cm^3}{mol}}$, $B_1 = -2455.297~\mathrm{\frac{K \cdot cm^3}{mol}}$, and $B_2 = -1.22675 \times 10^7~\mathrm{\frac{K^2 \cdot cm^3}{mol}}$.

The equilibrium pressure will be given by the value for which the function $\Delta G_{\text{sim}}(p) $, defined by Eq.~\eqref{eq:equilibrium pressure}, intersects $\Delta G_{\text{sim}} (p)$, as obtained in simulations. Note that besides  $\Delta G_{\text{sim}}$, we are also considering  variations of $c_s$ with pressure. These computations are greatly facilitated by the technique of reweighting \cite{McDonald1967,McDonald1969}, which uses simulation results from different states to improve the statistics and, particularly important to our case, allows to interpolate expectation values of thermodynamic observables along unsampled states. This implies a significant reduction in computational cost. In particular, we use the Multistate Bennett Acceptance Ratio (MBAR) estimator \cite{Shirts_2008}, which has been successfully used for the purpose of reweighing ~\cite{Aimoli2014, Nichele2018, Schieber2018}.

@@

The prediction of activity coefficients requires additional simulations of the solute as a pure liquid phase, whose properties we denote here by a superscript ``$0$''.
We start by using a standard thermodynamic relation \cite{Tester} to write the solute activity coefficient $\gamma_s$ in terms of chemical potentials as
\begin{equation}
\ln \gamma_s = \frac{\mu_s^l - \mu_s^{l,0}}{R T} - \ln x_s,
\end{equation}

@@

In the special case of a solute at infinite dilution in an ionic liquid solvent, its activity coefficient ($\gamma^\infty_s$) can be computed by replacing $\tilde{\rho}$ in the equation above by the molar density of the pure solvent, $\tilde{\rho}_\textit{IL}$.

@@

Finally, we also consider the parameterization Weber and Kirchner \cite{Weber_2016} carried out for \ce{[emim][B(CN)_4]} based on the force field developed by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
As the potential parameters for the anion were specified neither in Ref.~\citenum{Weber_2016} nor in Ref.~\citenum{Canongia_Lopes_2006}, we obtained them by direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure with Hartree-Fock calculations at the \mbox{6-31++G\textsuperscript{**}} level of theory, resulting in net charges of $\pm 0.8 ~\mathrm{e}$.

In the case of ionic liquid \ce{[emim][Tf_2N]}, we use the model of K\"{o}ddermann \textit{et al}. \cite{Koddermann_2007}, which was also derived by refining parameters of the force field by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
The authors characterized this refinement as unusual, given that they modified the LJ parameters while keeping the partial charges unchanged.
Kerl \textit{et al}. \cite{Kerl__2017} recently employed this model to accurately predict Henry constants of \ce{CO_2} in \ce{[emim][Tf_2N]} at several temperatures.

@@

With the aim of setting up a computationally efficient protocol for the IL simulations, in this work we assess the possibilities of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing electrostatic interactions.
In this method, the electrostatic potential between two atoms with charges $q_i$ and $q_j$ interacting at a distance $r$ is given by
\begin{equation}
V_\text{DSF}(r) = V(r) - V(r_c) - V^\prime(r_c)(r-r_c),
\end{equation}
where $V(r) = q_i q_j {\text{erfc}(\alpha r)}/{r}$ is a damped Coulomb potential.
This potential is cut off at $r = r_c$, which makes it short-ranged and thus a cheaper alternative to Ewald-type schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}.
McCann \textit{et al}. \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force field.
They concluded that using $\alpha = 0.2~\text{\AA}^{-1}$ and $r_c = 15~\text{\AA}$ provides the best energetic agreement with the Ewald method.
In this work, detailed comparisons are made in terms of the simulated densities and radial distribution functions.
In Sec.~\ref{sec:results}, we show that by using the same $\alpha$ value but a considerably smaller cutoff distance ($r_c = 12 \text{\AA}$) one can obtain suitable results.
These values were employed here for the ILs as well as for solutes containing partially charged atoms.
The only exception was hexane, for which the use of undamped Coulombic interactions (i.e. $\alpha = 0$) resulted in improved chemical potential predictions. That cutoff distance was also used for truncating the LJ interactions, for which we applied analytical tail corrections \cite{allen}. In addition, we used the Lorentz-Berthelot mixing rules to calculate the cross LJ interactions, except for the solutes modeled with the OPLS-AA force fields, which employs geometric combining rules.

The NpT algorithm of Kamberaj \textit{et al}. \cite{Kamberaj_2005} was used to integrate the equations of motion for rigid bodies.
For flexible molecules, we applied the integration scheme of Shinoda \textit{et al}. \cite{Shinoda2004}, with the lengths of all C-H bonds constrained by the SHAKE method \cite{Ryckaert1977}.
For both integrators, we used a small time-step size ($\Delta t$) of 1 fs in order to avoid discretization-induced artifacts \cite{Eastwood_2010, Silveira_2017}.

@@

In this work, we compute $\Delta G_\text{Coul}$ by carrying out an independent simulation of 10 ns at each state of $\lambda_{\text{Coul}}$.
In the end, we apply the MBAR method \cite{Shirts_2008} by using the \textit{Alchemical Analysis} package \cite{Klimovich_2015}.

In the case of $\Delta G_\text{vdW}$, the Expanded Ensemble method \cite{Lyubartsev_1992} is applied for sampling the whole $\lambda_\text{vdW}$ set in a single simulation.
An expanded ensemble is constructed as a weighted sum of subensembles, each one assuming a different value of the parameter selected for exploring the free energy landscape, which in our case is $\lambda_\text{vdW}$.
In the method adopted here \cite{Chodera_2011_2}, simulating an expanded ensemble comprises doing molecular dynamics (MD) at fixed N, p, T, and $\lambda_\text{vdW}$ amid Monte Carlo (MC) transitions in the value of $\lambda_\text{vdW}$.
The number of MD steps between transitions is fixed and, at each transition, one of the $\lambda_\text{vdW}$ states is picked at random according to a corresponding set of probabilities \cite{Chodera_2011_2}.
In addition, a set of biasing weights is included in the MC acceptance criterion, thus making it possible to enhance the sampling procedure by iteratively tuning these weights until all $\lambda_\text{vdW}$ states are visited with equal frequency \cite{Lyubartsev_1992}.
Given that sample decorrelation is improved by having short MD runs between MC transitions \cite{Chodera_2011_2}, we used only 5 steps in each run.
In the iteration process for tuning the biasing weights, the simulation times for the preliminary and the final expanded ensemble rounds are 10 and 40 ns, respectively.
In this way, 3 iterations were sufficient in all cases for attaining, in the final round, an acceptable uncertainty in $\Delta G_\text{sim}$ estimated via MBAR \cite{Shirts_2008, Klimovich_2015}. Considering also the MD runs at each $\lambda_{\text{Coul}}$, the simulation time was 110 ns for each system.

In order to obtain minimum-energy conformations, we first carried out MD simulations of isolated ion pairs.
For this, we used a Nos\'{e}-Hoover chain thermostat \cite{Martyna1992} with set-point temperature gradually changing down to nearly $0~\text{K}$.
The resulting coordinates were then used with the PACKMOL algorithm \cite{Mart_nez_2009} via the Playmol package \cite{playmol} to obtain initial liquid-phase configurations.
The expanded ensemble procedure, the softcore potential, and the DSF method  with linear coupling were implemented by customizing the LAMMPS package \cite{Plimpton1995}, which we used for carrying out all the simulations in this work.

@@

\subfloat[Liu~\textit{et al}.~\cite{Liu_2014}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Liu}}%
\subfloat[Weber and Kirchner~\cite{Weber_2016}]{%
  \includegraphics[width=.4\linewidth,trim={0 0 0 12pt},clip]{Weber}}%
\caption{Combined distribution functions of the dihedral angles corresponding to atoms C1-N2-C5-C6 and C2-N2-C5-C6 in the cation \ce{[emim]^+} (see Fig.~\ref{fig:atoms_id}), as sampled in NpT-MD simulations of 250 ion pairs using each model for \ce{[emim][B(CN)_4]}.}
\label{fig:dihedrals_emim}
\end{figure*}
